---
title: "Data Manipulation"
author: "Marisa Sobel"
date: "9/20/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

## Import FAS csv files 

Import first csv. (`FAS_litters.csv`). 

```{r}

litters_data = read_csv("./data/FAS_litters.csv",
  col_types = "ccddiiii")
litters_data = janitor::clean_names(litters_data)

pups_data = read_csv("./data/FAS_pups.csv",
  col_types = "ciiiii")
pups_data = janitor::clean_names(pups_data)

```

## Select variables 
Selecting columns 

name by name

```{r}
select(litters_data, group, litter_number, gd0_weight)
select(litters_data, litter_number, gd0_weight, pups_survive)
```

specific range:

```{r}
select(litters_data, gd_of_birth:pups_survive)
```

what you DON'T want to keep

```{r}
select(litters_data, -litter_number)
```

renaming while selecting 
just renaming if only changing 1 or 2 var, not have to type in all variables 

```{r}
select(litters_data, group, litter_number, gest_day_0_weight = gd0_weight)
rename(litters_data, gest_day_0_weight = gd0_weight)
```

use select helpers

```{r}
select(litters_data, starts_with("gd"))
select(litters_data, litter_number, starts_with("pup"))

# puts litters_number first
select(litters_data, litter_number, everything())  
```

### Practice
In the pups data, select the columns containing litter number, sex, and PD ears.

```{r}
select(pups_data, litter_number, sex, pd_ears)
```

## Filter obs

Eliminate some rows from dataset. 
Which are the rows you want to take out? 
_Identify a feature of some variable, TRUE=keep, FALSE=discard_

Do some filtering 

```{r}
filter(litters_data, gd0_weight < 25)
filter(litters_data, gd0_weight >= 25)

filter(litters_data, pups_born_alive == 8)
filter(litters_data, !is.na(gd0_weight))

filter(litters_data, group == "Low7" | group == "Low8")
filter(litters_data, group %in% c("Low7", "Low8"))

filter(litters_data, pups_born_alive %in% 4:6)

filter(litters_data, pups_born_alive %in% 4:7 & !is.na(gd0_weight))
```

### Practice
_**In the pups data:**_
Filter to include only pups with sex 1
Filter to include only pups with PD walk less than 11 and sex 2

```{r}
filter(pups_data, sex == 1)
filter(pups_data, sex == 2, pd_walk < 11)
```

## Mutate

Add new variable, change group names to lower-case

```{r}
mutate(litters_data,
       wt_gain = gd18_weight - gd0_weight, 
       group = tolower(group)
)

# new variable, and alter it immediatly

mutate(litters_data, 
       wt_gain = gd18_weight - gd0_weight,
       wt_gain_squared = wt_gain^2)
```

### Practice
_**In the pups data:**_
Create a variable that subtracts 7 from PD pivot
Create a variable that is the sum of all the PD variables

```{r}
mutate(pups_data, 
       new_var = pd_pivot - 7)
```

## Arrange 

Arrange the data

```{r}
arrange(litters_data, gd0_weight)
arrange(litters_data, pups_born_alive, gd0_weight)
arrange(litters_data, desc(pups_born_alive), gd0_weight)
```

## Pipeline for `litters_data`



